---
title: "Assignment_2"
author: "Joseph"
date: "2024-07-11"
output:
  pdf_document: 
    latex_engine: lualatex
  html_document: default
---
``` {r setup, include=FALSE}
knitr:: opts_chunk$set(echo = TRUE)
library(tidyverse)
library(datasets)
#install.packages("ggplot2")
library(ggplot2)
```

## Load the "mtcars" dataset from the built-in datasets in R. Write an R script to perform the following tasks:
```{r}
### Calculate the average miles per gallon (mpg) for all cars in the dataset.
data <- mtcars
avg_miles_per_gallon <-  mean(data$mpg)

cat("Average miles per gallon(mpg) is ", avg_miles_per_gallon ,"\n")

### Create a bar plot to visualize the number of cylinders (cyl) for each car.
barplot(data$cyl, main="Bar plot", xlab = "Number of cylinder", ylab = "count", col="skyblue", border = "black", lwd=3)

```

## Write an R script to read a CSV file named "sales_data.csv" containing sales data for different products. Perform the following tasks:

```{r}

library(dplyr)
library(tidyverse)

# Calculate the total sales for each product category.
sales_data <-  read.csv("sales_data.csv")

total_sales_by_product <- sales_data  %>% group_by(Product) %>%  summarise(sum_total_sales = sum(Sales)
 ) 

print(total_sales_by_product)

# Generate a line plot to visualize the monthly sales trend for one of the product categories.
sales_data$Date <- as.Date(sales_data$Date)

print(sales_data)

## Filter data for the specified product category and adding new month column
filtered_data <- sales_data %>% 
mutate(Month = format(sales_data$Date, "%b")) %>%  filter(sales_data$Product == "Heater")

print(filtered_data)

## Line Plot
plot(factor(filtered_data$Month),filtered_data$Sales,
     type="l", # type "o" for points and lines
     col = "blue",
     xlab = "Product",
     ylab = "Total Sales",
     lwd=5,
     main = paste("Monthly Sales Trend"))

## Add a legend
legend("topright", legend = "Heater", col = "green", lty = 1, cex = 0.8)

```

## Write an R script to generate a scatter plot using random data. The scatter plot should have 100 points with x and y coordinates ranging from 1 to 100. Add labels and a title to the plot.

```{r}
set.seed(1)

x <- rnorm(100)
y <- rnorm(100)

plot(x,y , main="Scatter between x and y", xlab = "x axis", ylab = "y axis", col="violet", lwd=4)

```

## Load the "iris" dataset from the built-in datasets in R. 
## Create a box plot to visualize the distribution of petal widths (Petal.Width) for each species.

```{r}
## Calculate the average sepal length (Sepal.Length) for each species.

grouped_data <- iris %>% group_by(Species)%>% 
 summarise(avg_sep_length <- mean(Sepal.Length))

print(grouped_data)

### Create a box plot to visualize the distribution of petal widths (Petal.Width) for each species
library(ggplot2)
ggplot(iris, aes(x = Species, y=Petal.Width)) +
  geom_boxplot(fill="skyblue") +
  ggtitle("Boxplot : Petal Width by species")

```