---
title: "Midterm_Exam"
author: "Joseph"
date: "2024-07-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Q1. Create a dataframe in R named `student_data` with columns "Name," "Age," and "Grade."
# Add at least five rows of data to the dataframe. . 
```{r}
students_data <-  data.frame(Name= c("jo", "christy", "kambham","Ahavah","Jay"),
                         Age = c(32,6,30,4,31),
                         Grade = c("A","A+","B","B+","A"))

print(students_data)

#create a list named `subject_grades`the grades of a student in three different subjects
subject_grades <- list(Maths = c("A", "A+", "B", "B+"), 
                       Science = c("A", "A+", "B","B+"),
                      History = c("A", "A+", "B","B+"))
print(subject_grades)
```

## Q2. Define a vector `num_vector` with values from 1 to 5. Convert this vector into a 2x3 matrix
## named `num_matrix`. Print both the vector and the matrix.
```{r}

num_vector <- c(1:5)
print(num_vector)

num_matrix <- matrix(num_vector, nrow = 2 , ncol = 3)
print(num_matrix)
# Since num_vector has only 5 elements, the matrix function will recycle the elements to fill the matrix.(i.e 1 is here)
```

## Q3. Create a factor variable `weather` with levels "Sunny," "Cloudy," and "Rainy." 

```{r}

# By changing the value assigned to the factor then the print statement changes
weather <- factor("Rainy", levels = c("Sunny", "Cloudy", "Rainy"))

# Write aconditional statement that prints "Take an umbrella" if the weather is "Rainy" and 
# "Enjoy the day" otherwise.
if (weather == "Rainy") {
  print("Take an umbrella")
} else {
  print("Enjoy the day")
}

```

## Q4. Load any CSV file named "sales_data.csv" into an R dataframe. 

```{r}
library(dplyr)
# Display the first 6 rows of the dataframe.
sales_data <-  read.csv("sales_data.csv")

head(sales_data,6)

# Calculate and print the average sales value
avg_sales <- mean(sales_data$Sales)
cat("Average sales across products", avg_sales , "\n")

```

## Q5. Consider any dataframe having `employee_data`  with columns "Name," "Salary," and "Department." 

``` {r}
# Creating the employee_data dataframe
employee_data <- data.frame(
  Name = c("John", "Jane", "Doe", "Alice", "Bob"),
  Salary = c(40000, 60000, 55000, 35000, 58000),
  Department = c("HR", "IT", "Finance", "IT", "Marketing")
)

# Write an R script that identifies and prints the names of employees whose salary
# is greater than $50,000 and belong to the "Marketing" department.

filter_data_df <- employee_data[employee_data$Salary > 50000 & employee_data$Department=="Marketing",]

print(filter_data_df)
```

# Q6. Create a list named `fruit_list` with elements "Apple," "Cherry," and "Strawberry." 
```{r}
fruit_list <- list("Apple" ,"Cherry" , "Strawberry")

# Write a function `add_fruit` that takes a fruit name as an argument and adds it to the list.
add_fruit <-  function(fruit_name) {

  new_list <- append(fruit_list,fruit_name)
  print(new_list)
  
}

#Call the function to add "Grapes" to the list and print the updated list

add_grapes <- add_fruit("Grapes")

```